Data Analysis

### 2023-03-28 ms1_data_analysis.Rmd
SLA models for fungicide and herbivore addition.

# Get the delta data ready for th e given subset vars (comparison)

@@

```{r}
simple.sla <- gls(
  lrrtb ~ lsla.d*site + I(lsla.d^2)*site + richness + labu + lrrsla,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

@@

AIC(just.sla, simple.sla)
anova(just.sla, simple.sla) # simple SLA is much bette
tab_model(simple.sla) # But only the lrr SLA is significant.

@@

anova(simple.sla, full.sla.linear) # full SLA linear is much better that the one using SLA
tab_model(full.sla.linear)

@@

anova(full.sla.linear, full.sla.linear.noint)
# Removing interactions does not improve the model significantly
tab_model(full.sla.linear.noint)

@@

anova(full.sla.linear.noint, full.sla.linear.intslawc)
# No effect
tab_model(full.sla.linear.noint)

@@

anova(full.sla.linear.noint, full.sla.linear.intsla)
# Any for of iteraction is not improving the model
tab_model(full.sla.linear.noint) # Thus the simplest model seems to be best.

@@

anova(full.sla.linear.noint, full.sla.custom)

@@

# Withouth a site it is even better
anova(full.sla.linear.noint, full.sla.linear.noint.nosite)

# The linear model remains the best one.
plot(full.sla.linear.noint.nosite) # This plot looks best
```

@@

>>>>

# Get the delta data ready for th e given subset vars (comparison)
subsetVars <- c("c", "i")

@@

```{r}
# Model explorations
simple.sla <- gls(
  lrrtb ~ lsla.d*site + I(lsla.d^2)*site + richness + labu + lrrsla,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

@@

# AIC(just.sla, simple.sla)
# anova(just.sla, simple.sla) # simple SLA is much bette
# tab_model(simple.sla) # But only the lrr SLA is significant.

@@

# anova(simple.sla, full.sla.linear) # full SLA linear is much better that the one using SLA
# tab_model(full.sla.linear)

@@

# anova(full.sla.linear, full.sla.linear.noint)
# Removing interactions does not improve the model significantly
# tab_model(full.sla.linear.noint)

@@

# anova(full.sla.linear.noint, full.sla.linear.intslawc)
# No effect
# tab_model(full.sla.linear.noint)

@@

# anova(full.sla.linear.noint, full.sla.linear.intsla)
# Any for of iteraction is not improving the model
tab_model(full.sla.linear.noint) # Thus the simplest model seems to be best.

@@

# anova(full.sla.linear.noint, full.sla.custom)

@@

# Withouth a site it is even better
# anova(full.sla.linear.noint, full.sla.linear.noint.nosite)

# The linear model remains the best one.
# plot(full.sla.linear.noint.nosite) # This plot looks best
# tab_model(full.sla.linear.noint.nosite)

@@

```{r}
subsetVars <- c("c", "f")

source('code/get_data_ready_for_the_sla_analyses.R')

source('code/model_definitions.R')

# Are traits improving the null model

# Model selection.
# 1. Full model (quadratic, and site interactions) vs no traits
# anova(no.traits, full.model)
# anova(no.traits, full.model.noint)
# anova(no.traits, all.linear)
# anova(no.traits, all.linear.noint)

# 2. Individual traits

# SLA
# Check quadratic
# anova(no.traits, just.sla.quadratic.noint)
# Check linear
# anova(no.traits, just.sla.linear.noint)
# Check quadratic interation
# anova(no.traits, just.sla.quadratic)
# Check linear interaction
# anova(no.traits, just.sla.linear)

# WATER
# Check quadratic
# anova(no.traits, just.water.quadratic.noint)
# Check linear
# anova(no.traits, just.water.linear.noint)
# Check quadratic interation
# anova(no.traits, just.water.quadratic)
# Check linear interaction
# anova(no.traits, just.water.linear)

# HERBIVORY
# Check quadratic
# anova(no.traits, just.herb.quadratic.noint)
# Check linear
# anova(no.traits, just.herb.linear.noint)
# Check quadratic interation
# anova(no.traits, just.herb.quadratic)
# Check linear interaction
# anova(no.traits, just.herb.linear)

# 2. Model with all traits linear/quadratic, what combination?

# quadratic SLA, HERB
# anova(no.traits, q.sla.q.herb)
# anova(no.traits, q.sla.q.herb.noint)
# quadratic SLA, linear HERB
# anova(no.traits, q.sla.l.herb)
# anova(no.traits, q.sla.l.herb.noint)
# linear SLA, quadratic HERB
# anova(no.traits, l.sla.q.herb)
# anova(no.traits, l.sla.q.herb.noint)
# linear SLA, linear HERB
# anova(no.traits, l.sla.l.herb)
# anova(no.traits, l.sla.l.herb.noint)

# quadratic HERB, WATER
# anova(no.traits, q.herb.q.water)
# anova(no.traits, q.herb.q.water.noint)
# quadratic HERB, linear WATER
# anova(no.traits, q.herb.l.water)
# anova(no.traits, q.herb.l.water.noint)
# linear HERB, quadratic WATER
# anova(no.traits, l.herb.q.water)
# anova(no.traits, l.herb.q.water.noint)
# linear HERB, linear WATER
# anova(no.traits, l.herb.l.water)
# anova(no.traits, l.herb.l.water.noint)

# quadratic SLA, WATER
# anova(no.traits, q.sla.q.water)
# anova(no.traits, q.sla.q.water.noint)
# quadratic SLA, linear WATER
# anova(no.traits, q.sla.l.water)
# anova(no.traits, q.sla.l.water.noint)
# linear SLA, quadratic WATER
# anova(no.traits, l.sla.q.water)
# anova(no.traits, l.sla.q.water.noint)
# linear SLA, linear WATER
# anova(no.traits, l.sla.l.water)
# anova(no.traits, l.sla.l.water.noint)

# So the best model is the one without the traits!
tab_model(no.traits)

# Additional herbivores

```{r}
subsetVars <- c("c", "h")

source('code/get_data_ready_for_the_sla_analyses.R')

source('code/model_definitions.R')

# Are traits improving the null model

# Model selection.
# 1. Full model (quadratic, and site interactions) vs no traits
anova(no.traits, full.model) # full model is much better
anova(full.model, full.model.noint)

# interaction model is as good as the no interaction. Keep the model without the interaction

# 2. Is the quadratic relationship significant
# anova(full.model.noint, all.linear)            # AIC is lower for the full model noint
# anova(full.model.noint, all.linear.noint)      # AIC is lower for the full model noint

# Traits have nonlinear relationships with the delta biomass.

# 3. REMOVE SLA
# 3.1 anova(full.model.noint, q.herb.q.water.noint) # withouth sla is slightly better (AIC)
# 3.2 anova(full.model.noint, q.herb.q.water)
# 3.3 anova(full.model.noint, q.herb.l.water.noint) # Full model is than this !!!
# 3.4 anova(full.model.noint, q.herb.l.water)
# 3.5 anova(full.model.noint, l.herb.l.water.noint) # Full model still better
# 3.6 anova(full.model.noint, l.herb.l.water) #

# Including SLA did not mprove the model significantly
# tab_model(q.herb.q.water.noint)

# But if we were to start first by remving perc.herbivory?
# anova(full.model.noint, q.sla.q.water.noint) # Model without herbivory is better
# anova(full.model.noint, q.sla.q.water)       # Full model is better
# anova(q.sla.q.water, q.sla.q.water.noint)    # No interaction is better
# anova(full.model.noint, q.sla.l.water.noint)   # Full model is still better
# anova(full.model.noint, q.sla.l.water)         # Full model wins again
# anova(full.model.noint, l.sla.l.water.noint)   # And again by AIC criterium
# anova(full.model.noint, l.sla.l.water)         # And again

# But if we were to start first by remving water.content?
# anova(full.model.noint, q.sla.q.herb.noint) # Model with herbivory is much better
# anova(full.model.noint, q.sla.q.herb)       # Full model is better
# anova(q.sla.q.herb, q.sla.q.herb.noint)    # No interaction is better
# anova(full.model.noint, q.sla.l.herb.noint)   # Full model is still better
# anova(full.model.noint, q.sla.l.herb)         # Full model wins again
# anova(full.model.noint, l.sla.l.herb.noint)   # And again by AIC criterium
# anova(full.model.noint, l.sla.l.herb)         # And again

# The above suggests that models without hebivory and sla do better job. Water content is the key trait that helps predict the change in biomass in response to the treatment.

# SLA
# Check quadratic
# anova(full.model.noint, just.sla.quadratic.noint)
# Check linear
# anova(full.model.noint, just.sla.linear.noint)
# Check quadratic interation
# anova(full.model.noint, just.sla.quadratic)
# Check linear interaction
# anova(full.model.noint, just.sla.linear)

# WATER
# Check quadratic
anova(full.model.noint, just.water.quadratic.noint) # Model with just water content

# Is the quadratic form necessary
anova(just.water.quadratic.noint, just.water.linear.noint) # no difference but quadratic has lower AIC

# So the best model is the one without the traits!
tab_model(just.water.quadratic.noint)
